---
title: "STOR 665 HW 7"
author: "Brian N. White"
date: "4/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MEMSS)
library(lme4)
library(lattice)
```

### Problem 7

### (a)

Explain the data briefly.
```{r load data, examine structure}
#load data
ergoStool <- ergoStool

#examine the data structure
str(ergoStool)
```
```{r summary}
#data summary
summary(ergoStool)
```
Inspection of the contingency table below reveals that the factors 'Type' and 'Subject' are completely crossed (i.e. there is at least one observation for each combination of factor levels). Further, this is an unreplicated design.
```{r}
xtabs(~Type + Subject, ergoStool)
```
### (b)

```{r}
#determine the average effort for each level of the factor Subject
ergoStool %>%
  group_by(Subject) %>%
  summarize(avg_effort=mean(effort)) -> avg_effort_df

ergoStool %>% 
  mutate(avg_effort=rep(avg_effort_df$avg_effort, each=4, times=1)) %>%
  mutate(Subject=fct_reorder(Subject, avg_effort)) %>%
  ggplot(aes(x=Subject, y=effort, shape=Type, color=Type)) +
  geom_point() +
  geom_line(aes(x=Subject, y=effort, group=Type, color=Type, linetype=Type))

```

### (c)

A linear mixed model with random effects for Type and Subject is fit below. The standard deviations for the estimates corresponding to Type, Subject and residual variability are 1.332, 1.695, and 1.100, as discerned from the output below.
```{r}
#fit a model with random effects for Type and Subject via REML
summary(lmm_ergo <- lmer(effort ~ 1 + (1|Type) + (1|Subject), ergoStool))
```
### (d)

The model from part (c) is refit using maximum likelihood estimation. The standard deviations for the estimates corresponding to Type, Subject and residual variability are now 1.305, 1.505, and 1.101. 
```{r}
summary(lmm_ergo2 <- update(lmm_ergo, REML=FALSE))
```
### (e)

In the code-chunk below, the 95% prediction intervals corresponding to the random effects of the model from part (d) (i.e. the model fit via MLE) are generated. Based upon this plot, it is clear that stool type 1 (i.e. T1) outperforms the other types w.r.t. effort minimization.
```{r}
## Substitute the name of your fitted model for fm in the call to ranef)
dotplot(ranef(lmm_ergo2, which = "Type", postVar = TRUE), aspect = 0.2, strip = FALSE)
```

