---
title: "STOR 665 HW 7"
author: "Brian N. White"
date: "4/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MEMSS)
library(lme4)
library(lattice)
```

### Problem 7

### (a)

The data in the ergoStool dataframe comes from an ergometrics experiment. In this experiment nine subjects were asked to sit on four stools, each with a different design type, and rate the difficulty of rising. The experimenter opted to measure 'effort' via the Borg scale. This scale ranges from  6-20 with integer values (i.e. with greater values corresponding to greater perceived effort). Thus, there are 36 observations (i.e. 9 subjects x 4 stool types) and 3 variables. The variable 'effort' is numeric and the remaining two variables are factors with levels corresponding to stool types and particular subjects.

```{r load data, examine structure}
#load data
ergoStool <- MEMSS::ergoStool

#examine the data structure
str(ergoStool)
```
```{r summary}
#data summary
summary(ergoStool)
```
Inspection of the contingency table below reveals that the factors 'Type' and 'Subject' are completely crossed (i.e. there is at least one observation for each combination of factor levels). Further, this is an unreplicated design.

```{r}
xtabs(~Type + Subject, ergoStool)
```
### (b)

The requested plot is output by the code-chunk below. Note, there is a minor irrelevant discrepancy between this plot and the corresponding plot in the package tutorial (i.e. the paths connecting points with the same Type level are not the same).

```{r}
#determine the average effort for each level of the factor Subject
ergoStool %>%
  group_by(Subject) %>%
  summarize(avg_effort=mean(effort)) -> avg_effort_df

ergoStool %>% 
  mutate(avg_effort=rep(avg_effort_df$avg_effort, each=4, times=1)) %>%
  #re-order the levels of Subject by avg_effort
  mutate(Subject=fct_reorder(Subject, avg_effort)) %>%
  ggplot(aes(x=effort, y=Subject, shape=Type, color=Type)) +
  geom_point() +
  geom_line(aes(x=effort, y=Subject, group=Type, color=Type, linetype=Type))

```

### (c)

A linear mixed model with random effects for Type and Subject is fit below. The standard deviations for the estimates corresponding to Type, Subject and residual variability are 1.332, 1.695, and 1.100, as discerned from the summary output.

```{r}
library(lme4)

#fit a model with random effects for Type and Subject via REML
summary(lmm_ergo <- lmer(effort ~ 1 + (1|Type) + (1|Subject), ergoStool))
```
### (d)

The model from part (c) is refit using maximum likelihood estimation. The standard deviations for the estimates corresponding to Type, Subject and residual variability are now 1.305, 1.505, and 1.101 (i.e. a decrease, relative to  the model in part (c) except for the residual standard deviation). 

```{r}
summary(lmm_ergo2 <- update(lmm_ergo, REML=FALSE))
```
### (e)

In the code-chunk below, the 95% prediction intervals corresponding to the random effects of the model from part (d) (i.e. the model fit via MLE) are generated. Based upon this plot, it is clear that stool type 1 (i.e. T1) outperforms the other types w.r.t. effort minimization.

```{r}
## Substitute the name of your fitted model for fm in the call to ranef)
dotplot(ranef(lmm_ergo2, which = "Type", postVar = TRUE), aspect = 0.2, strip = FALSE)
```

### (f)

The significance of the random effect Type is assessed by comparing the model fit in Part (d) with a reduced model with only Subject as the random effect. Inspection of the anova output indicates a p-value of approximately zero. Thus, we reject the null-hypothesis $H_{0}:\sigma_{2}=0$. This suggests that the more complex model, lmm_ergo2, fits the observed data better.

```{r}
summary(lmm_ergo3 <- lmer(effort ~ 1 + (1|Subject), ergoStool, REML=FALSE))
anova(lmm_ergo3, lmm_ergo2)
```

### (g)

The model in question is fit below (i.e. 'Type' is treated as fixed and 'Subject' as random) via MLE. For the fixed-effect parameter estimates note the following: relative to the model fit in part (d) the intercept (TypeT1) has decreased. Further, while TypeT2, TypeT3 and TypeT4 are positive, their values are smaller than what is suggested by the mean effort for the corresponding stool types.

```{r fixed effect for Type, random for Subject}
lmm_ergo4 <- lmer(effort ~ 1 + Type + (1|Subject), ergoStool, REML = 0)
summary(lmm_ergo4)

ergoStool %>% group_by(Type) %>% summarise(mean_effort=mean(effort))
```

